---
title: "Class 15: Pertussis Mini Project"
author: "Josie (A11433761)"
format: pdf
---

## CDC tracks Pertussis cases [here](
https://www.cdc.gov/pertussis/php/surveillance/pertussis-cases-by-year.html?CDC_AAref_Val=https://www.cdc.gov/pertussis/surv-reporting/cases-by-year.html)

## Web Scrapping
We need to scrape this data for R. Install **datapasta**
```{r}
cdc<-data.frame(
                               year = c(1922L,1923L,1924L,1925L,1926L,
                                        1927L,1928L,1929L,1930L,1931L,1932L,
                                        1933L,1934L,1935L,1936L,1937L,1938L,
                                        1939L,1940L,1941L,1942L,1943L,
                                        1944L,1945L,1946L,1947L,1948L,1949L,
                                        1950L,1951L,1952L,1953L,1954L,1955L,
                                        1956L,1957L,1958L,1959L,1960L,
                                        1961L,1962L,1963L,1964L,1965L,1966L,
                                        1967L,1968L,1969L,1970L,1971L,1972L,
                                        1973L,1974L,1975L,1976L,1977L,1978L,
                                        1979L,1980L,1981L,1982L,1983L,
                                        1984L,1985L,1986L,1987L,1988L,1989L,
                                        1990L,1991L,1992L,1993L,1994L,1995L,
                                        1996L,1997L,1998L,1999L,2000L,
                                        2001L,2002L,2003L,2004L,2005L,2006L,
                                        2007L,2008L,2009L,2010L,2011L,2012L,
                                        2013L,2014L,2015L,2016L,2017L,2018L,
                                        2019L,2020L,2021L,2022L,2024L),
       cases= c(107473,164191,165418,152003,
                                        202210,181411,161799,197371,166914,
                                        172559,215343,179135,265269,180518,
                                        147237,214652,227319,103188,183866,
                                        222202,191383,191890,109873,133792,
                                        109860,156517,74715,69479,120718,68687,
                                        45030,37129,60886,62786,31732,28295,
                                        32148,40005,14809,11468,17749,
                                        17135,13005,6799,7717,9718,4810,3285,
                                        4249,3036,3287,1759,2402,1738,
                                        1010,2177,2063,1623,1730,1248,1895,
                                        2463,2276,3589,4195,2823,3450,4157,
                                        4570,2719,4083,6586,4617,5137,
                                        7796,6564,7405,7298,7867,7580,9771,
                                        11647,25827,25616,15632,10454,13278,
                                        16858,27550,18719,48277,28639,
                                        32971,20762,17972,18975,15609,18617,
                                        6124,2116,3044,23544)
     )
```
```{r}
cdc
```
Now plot using ggplot2:
Call ggplot, add `cdc` inside ggplot, add `aes`, x and y axes, and then geom_point() and geom_line()
```{r}
library(ggplot2)
baseplot<-ggplot(cdc)+ 
  aes(year,cases)+
  geom_point()+
  geom_line()


baseplot
```
Let's add the data of wP vaccine roll out, 1946 (blue)
Let's add the data of aP vaccine roll out, 1996 (red)
Now let's add 2020 and see what happened to cases that year (gray)
Now, let's see the first spike after aP roll out, 2004

```{r}
baseplot+
  theme_bw()+
  geom_vline(xintercept = 1946, col="blue")+
  geom_vline(xintercept = 1996, col="red")+
  geom_vline(xintercept = 2020, col="gray")+
  geom_vline(xintercept = 2004, col="green")
```

## CMI-PB(Computational Models of Immunity-Pertussis Boost)

This projects collects and makes freely available data about the immune response to Pertussis vaccination
Access this data via an API which returns JSON format (key:vale pairs)
We can use **jsonlite** package and it's `read_json()` function

```{r}
library(jsonlite)
subject<- read_json("subject", simplifyVector = TRUE)
head(subject)
```
```{r}
table(subject$infancy_vac)
```

```{r}
table(subject$race, subject$biological_sex)
```

```{r}
table(subject$dataset)
```

```{r}
specimen<-read_json("specimen.json", simplifyVector = TRUE)
ab_titer<-read_json("plasma_ab_titer.json", simplifyVector = TRUE)
```
```{r}
head(specimen)
```
```{r}
head(ab_titer)
```

We want to "join" these tables so we have all the info we need about a given antibody measurement

```{r}
library(dplyr)
meta<-inner_join(subject,specimen)
```

```{r}
head(meta)
```
One last joining of `meta` and `ab_titer`
```{r}
abdata<- inner_join(ab_titer,meta)
head(abdata)
```

```{r}
nrow(abdata)
```

```{r}
table(abdata$isotype)
```


```{r}
table(abdata$antigen)
```

Let's begin with IgG
```{r}
igg<- filter(abdata,isotype=="IgG")
head(igg)
```
Make a boxplot of IgG antigen levels-this will be a plot of MFI vs antigen

```{r}
igg_boxplot<-ggplot(igg)+
  aes(MFI_normalised, antigen, col=infancy_vac)+
  geom_boxplot()
igg_boxplot
```
Ideally, I would like to see how these ab levels change over time relative to the booster shot
```{r}
table(abdata$visit)
  
```

```{r}
igg_boxplot+
  facet_wrap(~infancy_vac)
```

```{r}
igg_boxplot+
  facet_wrap(~visit)
```


```{r}
abdata.21 <- abdata %>% filter(dataset == "2021_dataset")
abdata.21 %>% filter(isotype == "IgG",  antigen == "PT") %>% 

ggplot()+
  aes(x=planned_day_relative_to_boost, y=MFI_normalised, col=infancy_vac,group=subject_id)+
  geom_point()+
  geom_line()+
  geom_vline(xintercept=0, linetype="dashed") +
  geom_vline(xintercept=14, linetype="dashed") +
  labs(title="2021 dataset IgG PT",
       subtitle = "Dashed lines indicate day 0 (pre-boost) and 14 (apparent peak levels)")


```





