---
title: "Class 11 AlphaFold Part 2"
author: "Josie (A11433761)"
format: pdf
---

```{r}
library(bio3d)
id<-"1ake_A"
aa<-get.seq(id)
aa
```
```{r}
b<-blast.pdb(aa)
```
```{r}
hits<-plot(b)
hits
```

```{r}
attributes(hits)
```
Tops hits from our BLAST results
```{r}
hits$pdb.id
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
I have now downloaded all adk structures in the PDB database but viewing is a mess. We need to align and superpossed.
Install BiocManager package from CRAN in the R console. Then I can use `BiocManager :: install()` to install any bioconductor package.

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```


```{r}
pdbs
```


```{r}
## Principal Component Analysis
pc<-pca(pdbs)
plot(pc)
```
```{r}
plot(pc,pc.axes=c(1:2))
```
To examine in more detail what PC1 (or any PC) is capturing here we can plot the loadings or make a movie (trajectory) or moving along the PC

```{r}
mktrj(pc,pc=1, file="pc1.pdb")
```

## Generating your own structure predictions




