---
title: "Class 09: Halloween Mini-Project"
author: "Josie (A11433761)"
format: pdf
---
## 1: Import data
```{r}
candy_file<-"candy-data.csv"
candy<-read.csv(candy_file, row.names = 1)
head(candy)

```
```{r}
nrow(candy)
```

```{r}
sum(candy$fruity)
```

```{r}
dim(candy)
```

## 2: Favorite Candy

```{r}
candy["Dum Dums",]$winpercent
```
```{r}
library(dplyr)
candy|>
filter( rownames(candy)=="Dum Dums")|>
  select(winpercent)

```
```{r}
candy|>
filter( rownames(candy)=="Kit Kat")|>
  select(winpercent)
```
```{r}
candy|>
  filter(rownames(candy)%in% c("Kit Kat", "Tootsie Roll Snack Bars"))|>
  select(winpercent)
```
`%in%` operator useful for checking the intersection


Install "skimr" in console
```{r}
library("skimr")
skim(candy)
```
Q6:winpercent column is on a very different scale than the rest of the data. It will need to be scaled.

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)

ggplot(candy)+ aes(winpercent)+
  geom_histogram(bins=10)+
  theme_bw()
```
```{r}
summary(candy$winpercent)
```
Q11: Chocolate ranks higher than fruity candy
```{r}
inds<-candy$chocolate==1
candy[inds,]$winpercent
```
```{r}
chocolate.win<-candy|>
  filter(chocolate==1)|>
  select(winpercent)
```
```{r}
fruit.win<-candy|>
  filter(fruity==1)|>
  select(winpercent)
```
```{r}
t.test(fruit.win,chocolate.win)
```

## 3: Candy Rankings

Lowest ranking candies: Nik L Nip,Boston Baked Beans,Chiclets,Super Bubble,Jawbusters
Highest ranking candies: Reese's Peanut Butter cup,Reese's Miniatures,Twix,Kit Kat,Snickers


```{r}
head(candy[order(candy$winpercent),], n=5)
sort(candy$winpercent)
inds<-order(candy$winpercent)
head(candy[inds,])
```
```{r}
inds<-order(candy$winpercent,decreasing=T)
head(candy[inds,],5)
```


```{r}
ggplot(candy)+
  aes(x=winpercent, y=reorder(rownames(candy),winpercent))+
  geom_col()
```
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy)+
  aes(x=winpercent, y=reorder(rownames(candy),winpercent))+
  geom_col(fill=my_cols)

```
Worst chocolate ranked: Sixlets
Best fruity candy ranked: Starbursts


```{r}
#Finding Sour Patch Kids and coloring the bar

my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols[rownames(candy)=="Sour Patch Kids"] = "blue"
ggplot(candy)+
  aes(x=winpercent, y=reorder(rownames(candy),winpercent))+
  geom_col(fill=my_cols)
 
```
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


## 4:Taking a look at pricepercent

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 8)
```
## 5: Exploring the correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
## 6: Principal Component Analysis

```{r}
pca <- prcomp(candy,scale=TRUE )
summary(pca)
```
```{r}
plot(pca$x[,2],col=my_cols,pch=16)
```
ggplot version
```{r}
my_data <- cbind(candy, pca$x[,1:3])
p<-ggplot(my_data)+
  aes(x=PC1, y=PC2,
        size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```
```{r}
p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 20)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24: Fruity,hard and pluribus make sense to be correlated with each other because that is how that type of candy is usually seen. The PC1 reflects that. 
