---
title: "Class 5: Data Viz with ggplot"
author: "Josie (A11433761)"
format: gfm
---

**base R** `plot()`
```{r}
plot(cars)
```
**ggplot2**
Before I can use any add-on package like this, I must install it.
`install.packages("ggplot2")` command/function
Then to use the package, I need to load it with a `library(ggplot2)` call

```{r}
library(ggplot2)
ggplot(cars) +
  aes(x=speed, y=dist) + 
  geom_point()
```
For "simple" plots, base R will be much shorter than ggplot code.

Let's fit a model and show it on my plot.

```{r}

ggplot(cars) +
  aes(x=speed, y=dist) + 
  geom_point() + 
  geom_smooth()
```
**data:** data.frame with the numbers you want to plot
**aes:** -thetics mapping of your data columns to your plot
**geom:** `geom_point()`, `geom_line()`. `geom_col()`

```{r}
mtcars
ggplot(mtcars) +
  aes(x=mpg, y=disp, size=hp)+
  geom_point( color="blue") 
  
```

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) + 
  geom_point() + 
  geom_smooth(method="lm", se=FALSE)
  
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```


```{r}
nrow(genes)
ncol(genes)
table(genes$State)
table(genes$State)/nrow(genes)


```
```{r}
p<-ggplot(genes)+
  aes(x=Condition1,y=Condition2, col=State)+
  geom_point()
p + scale_colour_manual(values=c("blue","gray","red"))+
  labs(title="Gene Expression Changes Upon Drug Treatment", x="Control (no drug)", y="Drug Treatment")
```




